Project Notes

### 2020-04-19 notes/notes.tex
Notes: update 18/04/2020
Notes: ideas on MPST web service formalism for client-client interaction

\documentclass{article}
\usepackage[margin=1in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{amsmath}

@@

\item
Adapt EFSM API -- previously, I use a metadata dictionary
to store the protocol name and role name. Now that I need to
propagate the custom type information too, I have a new 
\texttt{Endpoint} API which holds the EFSM (without metadata)
and the protocol/role name is lifted as a property of the 
endpoint API. Jinja templates have been adapted with the
API change.
\end{itemize}

@@

@@

\subsection*{Receive transitions in React}
This is the current workflow -- need to verify whether the receive transition is safe?

>>>>

\documentclass{article}
\usepackage[margin=1in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{wasysym}
\usepackage{stmaryrd}

@@

\item
Adapt EFSM API -- previously, I use a metadata dictionary
to store the protocol name and role name. Now that I need to
propagate the custom type information too, I have a new 
\texttt{Endpoint} API which holds the EFSM (without metadata)
and the protocol/role name is lifted as a property of the 
endpoint API. Jinja templates have been adapted with the
API change.

\item
Update Jinja templates to render custom type import
statements in TypeScript.
\end{itemize}

\paragraph{19/04/2020}
\begin{itemize}
\item
Draft theory of MPST for webservices over websockets 
(on OneNote).
\end{itemize}

@@

\newpage

\section{MPST for Web Services}

\paragraph{Main Idea}
Protocols for webservices implemented over WebSockets
present additional constraints.

\begin{enumerate}
\item Exactly one server-side role.
\item Client-side roles only \textit{directly} communicate 
to the server-side role.
\end{enumerate}

\subsection*{Global types for webservices}

Recall $G$, the global types from multiparty session types:

\begin{align*}
G ::&= \texttt{end} \\
&\mid p \to q : \{ l_i: G_i \}_{i \in I} \\
&\mid \mu t. G \\
&\mid t
\end{align*}

We consider $G_{WS}$, global types for webservices over 
websockets, defined as a tuple of the original global type $G$, 
and a participant $p$ explicitly specified as the server role;
all other participants $q \in pt(G)$ are assumed to take
on browser roles:
\begin{align*}
G_{WS} ::= G \times p
\end{align*}

We also use the same process calculus defined before:

\begin{align*}
P ::&= 0 \mid \mu X. P \mid X \\
&\mid p \rhd \{ l_i: P_i \}_{i \in I} \\
&\mid p \lhd l. P \\
&\mid \texttt{if e then } P \texttt{ else } P\\\\
\mathcal{M} ::&= p :: P \\
&\mid \mathcal{M} \mid \mathcal{M}
\end{align*}

The typing judgement for a multiparty session with respect
to the webservice global type builds upon the \textsc{[MTy]}
rule.

\subsection*{Codegen}
\texttt{codegen :: $\mathcal{M}$ $\times$ Role 
$\to$ $\mathcal{M}$}

\subsubsection*{Main rules}

\begin{align*}
&codegen~(p :: q \lhd l_j. P 
\mid q :: p \rhd \{ l_i: Q_i \}_{i \in I} 
\mid s :: S 
\mid \mathcal{M})~s \\
=~&codegen~(p :: s \lhd l. P 
\mid q :: s \rhd \{l_i: Q_i\}_{i \in I}
\mid s :: p \rhd )\\
&where  \\
\dots
\end{align*}

\subsubsection*{Trivial rules}

\subsection*{Encoding}

\subsection*{Preservation}
Modifying the process through code generation preserves
well-typedness of the session.

\paragraph{To show:} let $G_{WS} = (G, p)$. 
If $\mathcal{M}$ is well-typed with respect to $G_{WS}$,
then \texttt{codegen($\mathcal{M},~p$)} is well-typed
with respect to $\llbracket G_{WS} \rrbracket$

\newpage

@@

\subsection*{[Done] Receive transitions in React}
This is the current workflow -- need to verify whether the receive transition is safe?
